---
title: "README"
author: "Rubén Vásquez Arredondo"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

[Custom foo description](#income-in-chile)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Trigger the garbage collector
gc()

# Clear the console
cat("\014")

# Clear the workspace
rm(list = ls())


options(scipen = 999)


#libraries

library(data.table)
library(ggplot2)
library(leaflet)
library(sf)
library(chilemapas)
library(sp)


#data
casen<-fread("data\\processed\\Casen_2022_processed.csv",stringsAsFactors = T)

casen<-casen[!is.na(raw_salary) &
           raw_salary > 0,]

chile <- casen[rep(1:.N,expr)]

chile[,region:= factor(region, levels = c(
  "Región de Arica y Parinacota",
  "Región de Tarapacá",
  "Región de Antofagasta",
  "Región de Atacama",
  "Región de Coquimbo",
  "Región de Valparaíso",
  "Región Metropolitana de Santiago",
  "Región del Libertador Gral. Bernardo O'Higgins",
  "Región del Maule",
  "Región de Ñuble",
  "Región del Biobío",
  "Región de La Araucanía",
  "Región de Los Ríos",
  "Región de Los Lagos",
  "Región de Aysén del Gral. Carlos Ibáñez del Campo",
  "Región de Magallanes y de la Antártica Chilena"))]

```
## Income prediction in Chile

In an ever-evolving world of data science and analytics, understanding and predicting income trends is crucial for economic and social policy development. This GitHub project offers a comprehensive exploration of income prediction for individuals working in Chile, utilizing data from the CASEN (Encuesta de Caracterización Socioeconómica Nacional) survey.

**Project Overview:**

Chile's CASEN survey is a valuable source of socioeconomic data, capturing a wide array of information about its population. This project delves into the realm of data science to harness this dataset's potential and provide insightful predictions on income levels for individuals across the country.

**Key Objectives:**

- **Data Exploration:** We start by delving deep into the CASEN data to understand the demographics, education levels, and other critical features that impact an individual's income.

- **Feature Engineering:** To improve model accuracy, we engineer and preprocess the dataset, ensuring it is ready for predictive modeling.

- **Machine Learning:** Employing advanced machine learning techniques, we build predictive models that can estimate an individual's income based on their socioeconomic characteristics.

- **Evaluation and Insights:** We assess the models' performance, interpret feature importance, and provide valuable insights into the factors that influence income disparities in Chile.

- **Visualization:** We employ data visualization techniques to make the results accessible and understandable, aiding not only data scientists but also policymakers, economists, and the general public in grasping the income dynamics in Chile.

**Why this Matters:**

Understanding income disparities is vital for designing targeted policies, reducing inequality, and enhancing economic well-being. The predictions and insights derived from this project have the potential to inform policymakers, researchers, and the public about the socioeconomic landscape in Chile.

# Income in Chile

The Casen Survey, short for "National Socioeconomic Characterization," is a periodic socioeconomic study conducted in Chile by the Ministry of Social Development and Family. It aims to collect relevant information on the economic and social situation of the Chilean population. It covers a wide range of topics, including demographics, education, health, housing, employment, income, and poverty, with data collected through random household interviews. In this case, we will use the 2022 version, the latest to the creation of this document. 

The CASEN recovered data of 202,231 people from all the regions of Chile (1% of the population of 19,828,563), from which 58,240 had some kind of remuneration in the last month. this subset is our target population. to understand them, lest see some basic statistics:

```{r variables, echo=FALSE, comment=""}

message(paste("The mean income in Chile is:",
            prettyNum(mean(chile$raw_salary),big.mark=",", digits = 5),
            "Chilean Pesos"
            )
      )


message(paste("The median income in Chile is:",
            prettyNum(median(chile$raw_salary),big.mark=",", digits = 5),
            "Chilean Pesos"
            )
      )


message(paste("The income`s standard deviation in Chile is:",
            prettyNum(sqrt(var(chile$raw_salary)),big.mark=",", digits = 5),
            "Chilean Pesos"
            )
      )


quantile(casen$raw_salary, probs = seq(0, 1, 0.1 ))


```
Wait a second. the standar desviation is bigger than the mean, that means that the distribution is skewed to the left. lest see the distribution:

```{r quantiles, echo=FALSE, comment=""}
message("The quantiles are the following:")
quantile(casen$raw_salary, probs = c(0.25,0.5,0.75,0.9,0.95,0.99,0.999,0.9999))

message("And the density plot:")

```

```{r distribution_plot, echo=FALSE}

ggplot(data = chile, aes(x = raw_salary)) +
  geom_density(fill = "skyblue", color = "black") +
  labs(title = "Income Distribution",
       x = "Income",
       y = "Frequency") +
  theme_minimal()


```
Not only the distribution is skewed to the left, but also the data is very concentrated in the lower quantiles. this is a problem because the model will be biased to predict low incomes. lets see the distribution of the income by region, from north to south:

```{r distribution_by_region, echo=FALSE,warning=FALSE}



ggplot(data = chile, aes(x = raw_salary , y = region)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Income Distribution by Region",
       x = "Income",
       y = "Region") +
  theme_minimal() +
  lims(x = c(0, 1000000)) +
  scale_y_discrete(limits=rev)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


```{r mapa, echo=FALSE}



```

```{r distribution_by_region_mapa, echo=FALSE,warning=FALSE, fig.show="hold", out.width="50%"}


```